UMD Building Classification

### 2024-12-09 report/paper/draft.tex
an error correction
revisions

\usepackage{graphicx} % Required for inserting images
\usepackage{tikz}
\usepackage{soul}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{float}
\usepackage{parskip}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage[shortlabels]{enumitem}

@@

Grad-CAM++ can be used as an important means for highlighting image classification confidence, both of which we intend to use as the most advanced methods in their respective lineages. As an example, see our preliminary original Grad-CAM results in Figure 1 on instances on our test set, where Grad-CAM indicates it's making predictions based on unrelated objects (foreground foliage, people, unrelated elements of the building). This can be used for confirmation of model confidence. If the model is sufficiently confident, this can even be used as a means to indicate building locations (especially given Grad-CAM++ ability to find multiple instances) within the image rather than using traditional segmentation techniques.

@@

Data was collected by individually taking pictures and labeling them, along with data augmentation. We took around 4000 photos of UMD buildings over 20 classes, aiming for a 60/40 split of daytime/nighttime photos.

@@

We achieved this using the pretrained ResNet18 model and fine-tuning it over our dataset. Code and parameters can be found in the github repository. The primary function of the FS model was to provide a baseline for our semi-supervised methods.

@@

@@

We trained a variety of different models with different hyperparameters (batch size, momentum, warm up epochs, pseudo label threshold, etc) and selected 5 for further testing. Parameters can be found on the github in \verb*|dataset/parms.ipynb|. All images were standardized at 224x224, using the SyRA synthesized rain test image set with a random augmentation pipeline.

All models were trained for 128 epochs over the WideResNet18 model with depth 16, widen factor 2, and dropout 1. Other architectures (within the ResNet family and VGG families) were tried as well. The model was trained on the 20 classes of UMD buildings, on Google Colab with a Tesla T4 GPU. All code is available on the github repository.

@@

All data went to both of our models, where it was then augmented with random horizontal translations and translations (eg. weak augmentation). We utilized MSE loss for consistency loss in unlabeled data. For labeled data, we utilized the weights formula \begin{align*}
    \theta_{\text{teacher}}=\alpha \cdot\theta_{\text{teacher}}+(1-\alpha)\cdot\theta_{\text{student}}
\end{align*} where $\alpha=0.97$ is the decay rate. Final loss was calculated as \begin{align*}
    \mathcal{L}_{\text{total}}=\mathcal{L}_{\text{supervised}}+\lambda\cdot\mathcal{L}_{\text{consistency}}
\end{align*}where $\lambda=1.75$ was the weight of the consistency loss.

The model was trained for 25 epochs over the WideResNet18 model. The model was trained on the 20 classes of UMD buildings, on Google Colab with a Tesla T4 GPU. All code is available on the github repository.

@@

VQA was found to be infeasible due to time constraints; the amount of questions required was too high given our priorities on semi supervised learning and the lack of time we allotted to the experiment.
% results for these will go into appendix

@@

The SLIC transform results in significant preprocessing costs, and was the main cost of the model. All code is available on the github repository, although no results are provided.

@@

We trained various ResNet/VGG models for the fully supervised baseline. After training, ResNet18 was determined to be the best model for semisupervised with a 97.63\% validation accuracy on the test set after 12 epochs. All training was done on the 20 classes of UMD buildings, on Google Colab with a Tesla T4 GPU. All code is available on the github repository.

All testing was done on the heavily augmented test set to check for robustness, and the table below was generated at 10 epochs.

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Architecture & Val Accuracy & Test Accuracy \\
        \hline
        DenseNet121 & 97.47\% & 87.77\% \\
        \hline
        DenseNet201 & 95.26\% & 87.93\% \\
        \hline
        ResNet18 & 97.63\% & 88.54\% \\
        \hline
        ResNet34 & 98.89\% & 86.38\% \\
        \hline
        ResNet50 & 96.05\% & 80.13\% \\
        \hline
    \end{tabular}
    \caption{Comparison of DenseNet/ResNet Model Accuracies}
    \label{tab:dnrn_model_accuracies}
\end{table}

@@

However, we were successful in locking Grad-CAM onto building features, and we achieved reasonable test accuracies on the heavily augmented SyRA test set.

@@

Our MeanTeacher model achieved top1 student accuracy of 75.08\% and top1 teacher accuracy of 73.89\%. This utilized $\lambda=1.5, \alpha=.99$, with batch size $0.001$, SGD decay $10^{-4}$, and momentum $.9$ over 20 epochs.

@@

Future research will focus on increasing the diversity of the dataset, along with continuing preliminary work on GNNs.

@@

>>>>

% All packages added by Jeffery
\usepackage{graphicx} % Required for inserting images
\usepackage{tikz}
\usepackage{soul}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{float}
\usepackage{parskip}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage[shortlabels]{enumitem}
\usepackage{subcaption}

@@

Grad-CAM++ can be used as an important means for highlighting image classification confidence, both of which we intend to use as the most advanced methods in their respective lineages. As an example, see our Grad-CAM results in Appendix \ref{extra_figures} on instances on our test set. This can be used for confirmation of model confidence. If the model is sufficiently confident, this can even be used as a means to indicate building locations (especially given Grad-CAM++ ability to find multiple instances) within the image rather than using traditional segmentation techniques.

@@

Data was collected by individually taking pictures and labeling them, along with data augmentation. We took around 4000 photos of UMD buildings over 20 classes, aiming for a 60/40 split of daytime/nighttime photos. Buildings included a diverse set of buildings but also included 4 dorms of extremely similar construction.

@@

We achieved this using the pretrained ResNet18 model and fine-tuning it over our dataset. Code and parameters can be found in the Github repository. The primary function of the FS model was to provide a baseline for our semi-supervised methods.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{model_architecture.png}
    \caption{Grad-CAM Model Architecture}
    \label{fig:gradcam_architecture}
\end{figure}

@@

We trained a variety of different models with different hyperparameters (batch size, momentum, warm up epochs, pseudo label threshold, etc) and selected 5 for further testing. Parameters can be found on the github in \verb*|CMSC472_Final_Project/code/my_fixmatch/dataset/parms.ipynb|. All images were standardized at 224x224, using the SyRA synthesized rain test image set with a random augmentation pipeline.

Models 1, 2, 3 were trained for 128 epochs, and 4, 5 for 32 over the WideResNet model with depth 16, widen factor 2, and dropout 1. Each model was trained on the previous, iteratively. The model was trained on the 20 classes of UMD buildings, on a NVIDIA RTX 3090. All code is available on the Github repository.

@@

All data went to both of our models, where it was then augmented with random horizontal translations and translations (eg. weak augmentation). We utilized MSE loss for consistency loss in unlabeled data. For labeled data, we utilized the weights formula \begin{align*}
    \theta_{\text{teacher}}=\alpha \cdot\theta_{\text{teacher}}+(1-\alpha)\cdot\theta_{\text{student}}
\end{align*} where $\alpha$ is the decay rate. Final loss was calculated as \begin{align*}
    \mathcal{L}_{\text{total}}=\mathcal{L}_{\text{supervised}}+\lambda\cdot\mathcal{L}_{\text{consistency}}
\end{align*}where $\lambda$ was the weight of the consistency loss.

The model was trained for 25 epochs over the ResNet18 model. The model was trained on the 20 classes of UMD buildings, on Google Colab with a Tesla T4 GPU. All code is available on the Github repository.

@@

VQA was found to be infeasible due to time constraints; the amount of questions required was too high given our priorities on semi-supervised learning and the lack of time we allotted to the experiment.
% results for these will go into appendix

@@

The SLIC transform results in significant preprocessing costs, and was the main cost of the model. All code is available on the Github repository, although no results are provided.

@@

We trained various ResNet/VGG models for the fully supervised baseline. After training, ResNet18 was determined to be the best model for semisupervised with a 97.63\% validation accuracy on the test set after 12 epochs. All training was done on the 20 classes of UMD buildings, on Google Colab with a Tesla T4 GPU. All code is available on the Github repository.

All testing was done on the heavily augmented test set to check for robustness, and we achieved the following peak accuracies for the models within 25 epochs:

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Architecture & Best Val Accuracy & Test Accuracy \\
        \hline
        DenseNet121 & 98.26\% & 87.77\% \\
        \hline
        DenseNet201 & 97.00\% & 87.93\% \\
        \hline
        ResNet18 & 97.63\% & 88.54\% \\
        \hline
        ResNet34 & 98.89\% & 86.38\% \\
        \hline
        ResNet50 & 96.05\% & 80.13\% \\
        \hline
    \end{tabular}
    \caption{Comparison of DenseNet/ResNet Model Accuracies}
    \label{tab:dnrn_model_accuracies}
\end{table}

@@

However, we were successful in locking Grad-CAM onto building features, and we achieved reasonable test accuracies on the heavily augmented SyRA test set. If time permits, we would like to add more labeled data to the dataset and see if the result improves.

@@

Our MeanTeacher model achieved top1 student accuracy of 75.08\% and top1 teacher accuracy of 73.89\%. This utilized $\lambda=1.5, \alpha=.99$, with $lr=0.001$, batch size $128$, SGD decay $10^{-4}$, and momentum $.9$ over 20 epochs.

@@

Future research will focus on increasing the diversity of the dataset, along with continuing preliminary work on GNNs. We would also like to experiment with more fine tuning of parameters and warm up functions.

@@

\begin{figure}[H]
    \centering
    % First subfigure
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{new_gradcam_result2.png}
        \caption{Grad-CAM visualization on ResNet18}
        \label{fig:gradcam1}
    \end{subfigure}
    \hfill  % Adds horizontal spacing
    % Second subfigure
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{new_gradcam_plusplus_result2.png}
        \caption{Grad-CAM++ visualization on ResNet18}
        \label{fig:gradcam2}
    \end{subfigure}

    \caption{Comparison of Grad-CAM and Grad-CAM++ visualizations on ResNet18}
    \label{fig:gradcam_comparison1}
\end{figure}

\begin{figure}[H]
    \centering
    % First subfigure
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{gradcam_result3.png}
        \caption{Grad-CAM visualization on MeanTeach}
        \label{fig:gradcam3}
    \end{subfigure}
    \hfill  % Adds horizontal spacing
    % Second subfigure
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mean_teach_gradcam_plusplus.png}
        \caption{Grad-CAM++ visualization on MeanTeach}
        \label{fig:gradcam4}
    \end{subfigure}

    \caption{Comparison of Grad-CAM and Grad-CAM++ visualizations on MeanTeach}
    \label{fig:gradcam_comparison2}
\end{figure}
